%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% abstract-en.tex
%% NOVA thesis document file
%%
%% Abstract in English
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE abstract-en.tex}%

Cloud computing has become a cornerstone of modern software systems, offering organizations unparalleled flexibility, scalability and cost efficiency. As businesses increasingly shift from on-premises infrastructures to cloud-native architectures, traditional desktop applications face growing limitations in terms of accessibility, maintainability and integration with distributed services. Web-based solutions, supported by elastic cloud platforms, enable continuous availability, improved security practices and simplified deployment pipelines. This paradigm shift has established cloud computing not only as a technological trend but as a standard model for delivering reliable and user-centric software at scale.

This thesis presents the design and migration of \sysname, a Java 8 desktop application originally built for local network environments, to a modern cloud-hosted web application. \sysname{} is a platform responsible for issuing invoicing documents, generating and communicating Standard Audit File for Tax (Portugal) files to the Portuguese Tax Authority, submitting Stamp Duty declarations and integrating invoicing data with banking and Customer Relationship Management systems through web services. The work includes defining a scalable and secure cloud-based architecture, redesigning the system into a modular web application and implementing a proof of concept that validates the proposed approach. This migration addresses the challenges of adapting a locally deployed, desktop-based solution to a distributed, service-oriented environment while preserving the robustness and reliability required for its operational context. The resulting solution demonstrates how legacy desktop applications can be successfully transformed to meet the demands of modern cloud-enabled ecosystems.

% Abstract keywords
\keywords{
  Cloud Computing \and
  Web Application Migration \and
  Software Architecture \and
  Cloud Deployment
}
